Scalar_M1

### 2015-07-20 models/Model_Playspace.Rmd
fiddling with model

```{r}
priors = read.csv("models/prior.csv")
# Possible transformations --->
priors$prior.p = sqrt(priors$prior.p)/(sum(sqrt(priors$prior.p)))
#priors$prior.p = (1/(-log(priors$prior.p))) / sum((1/(-log(priors$prior.p))))
#priors$prior.p = (log(priors$prior.p + 1)) / sum((log(priors$prior.p + 1)))
speaker = read.csv("models/speaker.csv")
#speaker.alt = read.csv("models/e5a.csv")
listener = read.csv("models/listener.csv")
```

@@

```{r}
#cost function in length of chars
# u.cost = function(u) {
#   return(nchar(u))
# }

#speaker likelihood helper 
speaker.lhd = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost)) #Rethink alpha and costs?
}

#speaker likelihood normalized
speaker.prob = function(rating, degree, m, alpha) {
  num = speaker.lhd(m[rating, degree], alpha)
  norm = 0
  for (i in 1:ncol(m)) {
    norm = norm + speaker.lhd(m[rating, i], alpha)
  }
  return(num / norm)
}

#non-normalized posterior
nn.post = function(rating, degree, m, alpha, useprior) {

@@

#normalized posterior
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(i, degree, m, alpha, useprior)
  }
  return(nn / norm)
}  
```

@@

```{r}
run.rsa <- function(d, alpha = 1, useprior = TRUE, usenone = FALSE) {

@@

```{r}
data <- left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

@@

```{r}
# Which alpha has tightest correlation with human judgements
alphas = seq(from=1, to=20)
fit = sapply(alphas, FUN=function(n) {
  md <- data %>%
    do(run.rsa(., alpha = n, useprior = TRUE, usenone = TRUE))

  mean((md$pred - md$listener.p)^2)
  #cor(md$pred, md$listener.p)
})
qplot(alphas, fit, main="MSE for alpha values", ylab="MSE")
best.alpha = which(fit == min(fit))
best.alpha

# RUN MODEL
md <- data %>%
  do(run.rsa(., alpha = best.alpha, useprior = TRUE, usenone = TRUE))
```

@@

```{r}
# Facets all scales, Model vs Human judgements
qplot(stars, listener.p, col=degree, 
      data=md, main="Model vs Human judgments",
      ylab="Posterior judgments p(r|u)") + 
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

ggplot(md, aes(x=listener.p, y=pred)) +
    geom_point(aes(colour = stars)) +
    geom_smooth(method=lm) + ggtitle("Model pred vs Human judgment")
cor(md$pred, md$listener.p)
```

@@

# Scalar compatibility (information ratings)
qplot(x = stars, y = speaker.p, col = degree, ymax = 1,
      geom = c("point", "smooth"), stat = "identity", 
      position = "dodge", 
      data = speaker, main= "Scalar Compaibility distr") + 
  facet_wrap(~scale)
```

Pragmatic Listener Judgments
```{r}
qplot(x=stars, y=listener.p, col=degree, ymax=1,
      geom=c("point", "smooth"), stat = "identity",
      ylab="Compatibility", position="dodge", 
      data=listener, main="Pragmatic listener judgements") + 
  facet_wrap(~scale)
```

@@

Priors
```{r}
qplot(c(1:5), priors$prior.p,
      geom="bar", stat = "identity",
      binwidth = 1, main="Priors over ratings")

shapiro.test(priors$prior.p)
ks.test(priors$prior.p, "pnorm", mean=mean(priors$prior.p), sd=sd(priors$prior.p))
```

>>>>

```{r}
# Priors ---------------->
emp.priors = read.csv("models/prior.csv")
yelp.priors = read.csv("models/yelp1_priors.csv")
priors = yelp.priors
# Possible transformations --->
#priors$prior.p = sqrt(priors$prior.p)/(sum(sqrt(priors$prior.p)))
#priors$prior.p = (1/(-log(priors$prior.p))) / sum((1/(-log(priors$prior.p))))
#priors$prior.p = (log(priors$prior.p + 1)) / sum((log(priors$prior.p + 1)))

# Speaker informativity ------------------------->
speaker = read.csv("models/speaker.csv")
speaker = read.csv("models/speaker_normed.csv") # for printing out semantic distr's
speaker = read.csv("models/yelp1.csv")
#speaker = read.csv("models/e5a.csv")

# Listener judgments ---------------------------->
listener = read.csv("models/listener.csv")
```

@@

```{r}
# cost 
# ----
# u.cost = function(u) {
#   return(nchar(u))
# }

# speaker likelihood helper 
# ------------------------
speaker.lhd = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost)) #Rethink alpha and costs?
}

# speaker likelihood normalized
# -----------------------------
speaker.prob = function(rating, degree, m, alpha) {
  num = speaker.lhd(m[rating, degree], alpha)
  norm = 0
  for (i in 1:ncol(m)) {
    norm = norm + speaker.lhd(m[rating, i], alpha)
  }
  return(num / norm)
}

# non-normalized posterior 
# ------------------------
nn.post = function(rating, degree, m, alpha, useprior) {

@@

# normalized posterior
# --------------------
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(i, degree, m, alpha, useprior)
  }
  return(nn / norm)
}  
```

@@

```{r}
# Run rsa model
# --------------------
run.rsa <- function(d, alpha = 1, useprior = TRUE, usenone = FALSE) {

@@

```{r}
# data
# ----
data <- left_join(speaker, listener) %>%
  left_join(emp.priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

@@

```{r}
# Tuning alpha, experimenting with params
# ---------------------------------------

# Which alpha has tightest correlation with human judgements
alphas = seq(from=1, to=20)
fit = sapply(alphas, FUN=function(n) {
  md <- data %>%
    do(run.rsa(., alpha = n, useprior = T, usenone = TRUE))

  #return(mean((md$pred - md$listener.p)^2))
  return(cor(md$pred, md$listener.p))
})
qplot(alphas, fit, main="Cor for alpha values", ylab="MSE")
best.alpha = which(fit == max(fit))
best.alpha

# RUN MODEL
# ---------
md <- data %>%
  do(run.rsa(., alpha = best.alpha, useprior = T, usenone = TRUE))
```

@@

```{r}
# Facets all scales, Model vs Human judgements
qplot(stars, listener.p, col=degree, 
      data=md, main="Model vs Human judgments\nSemantics measured empirically",
      ylab="Posteriors p(r|u)") + 
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

ggplot(md, aes(x=pred, y=listener.p)) +
    geom_point(aes(colour = stars)) +
    geom_smooth(method=lm) +
  ggtitle("Model performance\nCorpus based") +
  labs(x = "model prediction", y="human judgments")
cor(md$pred, md$listener.p)

# Individual scales ---------->
# Scalar compatibility (information ratings) - individual plot
current_scale = subset(md, scale=="good_excellent",
                    select=c(stars, scale, speaker.p, listener.p, degree, pred))
levels(current_scale$degree) = c("excellent", "good")
qplot(stars, listener.p, col=degree, 
      data=current_scale, main="Model vs Human judgments\nCorpus based",
      ylab="p(stars|scalar_term)") +
geom_line(aes(y = pred), lty = 4)

```

@@

# Scalar compatibility (information ratings) - facet plot
qplot(x = stars, y = speaker.normed, col = degree, ymax = 1,
      geom = c("point", "smooth"), stat = "identity", 
      position = "dodge", ylab="P(r|u)",
      data = speaker, main= "Scalar compatibility\nCorpus based") + 
  facet_wrap(~scale)

# Scalar compatibility (information ratings) - individual plot
current_scale = subset(speaker, scale=="good_excellent",
                    select=c(stars, scale, speaker.p, degree))

levels(current_scale$degree) = c("excellent", "good")
qplot(x = stars, y = speaker.p, col = degree, ymax = .75,
      geom = c("point", "smooth"), stat = "identity", 
      position = "dodge", ylab="p(stars|scalar_term)",
      data = current_scale,
      main= "Literal listener semantics\nCorpus data")
```

Pragmatic Listener Judgments
```{r}
qplot(x=stars, y=listener.p, col=degree, ymax=1,
      geom=c("point", "smooth"), stat = "identity",
      ylab="Compatibility", position="dodge", 
      data=listener, main="Pragmatic listener judgements") + 
  facet_wrap(~scale)

# Individual scales ---------->
current_scale = subset(listener, scale=="good_excellent",
                    select=c(stars, scale, listener.p, degree, pred))
levels(current_scale$degree) = c("excellent", "good")
qplot(stars, listener.p, col=degree, xmin=0,
      geom = c("point", "smooth"), stat = "identity", 
      data=current_scale, main="Pragmatic listener",
      ylab="p(stars|scalar_term)")
```

@@

Priors
```{r}
qplot(c(1:5), yelp.priors$prior.p,
      geom="bar", stat = "identity",
      binwidth = 1, xlab="stars", ylab="p(stars)",
      main="Priors\nCorpus based")

shapiro.test(emp.priors$prior.p)
ks.test(emp.priors$prior.p, "pnorm",
        mean=mean(emp.priors$prior.p), sd=sd(emp.priors$prior.p))
```
